---
title: "10_Load_Data"
output: html_notebook
---

# What affects the number of daily fligths?

## First, load in important libraries

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(modelr)
```

## Load in flights and related data

```{r}
library(readr)
flights <- read_csv(
  "data/flights.csv",
  col_types = cols(
    arr_delay = col_integer(),
    arr_time = col_integer(),
    day = col_integer(),
    dep_delay = col_integer(),
    dep_time = col_integer(),
    flight = col_character(),
    month = col_integer(),
    sched_arr_time = col_integer(),
    sched_dep_time = col_integer(),
    year = col_integer()
  )
)
```

## What is the flight volume by day?

There is no well formatted date time field for the departures. I will need to create a new column to be used throughout the analysis which mutates the current encoding.

```{r}
daily <- flights %>%
  mutate(date = make_date(year, month, day)) %>%
  group_by(date) %>%
  summarise(count = n() ) %>%
  arrange(desc(count))
daily
```

Busiest day was day before thanksgiving

```{r}
daily <- flights %>%
  mutate(date = make_date(year, month, day)) %>%
  group_by(date) %>%
  summarise(count = n() ) %>%
  arrange(count)
daily
```

Least traveled day was day OF thanksgiving

Let's graph the time series of number of flights

```{r}
ggplot(data = daily) + 
  geom_line(mapping = aes(x = date, y = count))
```

Spikes, drops in volume might be day of week
We need an analysis of the dirtribution of flights by day of week

We need to indicate day of week in data frame

```{r}
daily <- daily %>%
  mutate(wday = wday(date, label = TRUE))
daily
```

Now plot distribution of flights by day of week

```{r}
daily %>%
  ggplot(mapping = aes(wday, count)) +
  geom_boxplot()
```

I want to remove the efect of day of week. I will fit a linear regression model and graph the residuals.

```{r}
mod <- lm(count ~wday, data = daily)

daily <- daily %>%
  add_residuals(mod)
daily
```

Now plot the residuals.

```{r}
daily %>%
  ggplot(mapping = aes(date, resid)) +
  geom_line()
```

drop in travel in january
increase in travel in summer




